Responsible Machine Learning Techniques\\\large{Interpretable Models, Post-hoc Explanation, and Discrimination Testing}

### 2019-12-18 article-information-2019.tex
more hdma results text

@@

Results for the mortgage data are presented in Subsections \ref{ssec: c_v_uc_mort} -- \ref{ssec:dis_mort}.  $g^\text{ANN}$ and $g^\text{XNN}$ outperform $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data, but as in Subsection \ref{ssec: c_v_uc_sim} the constrained variants of both model architectures do not show large differences in model performance with respect to unconstrained variants. Assuming that small fit differences do not outweigh the need for intrinsic model interpretability and post-hoc explainability in high-stakes, human-centered, or regulated applications, $g^\text{MGBM}$ and $g^\text{XNN}$ interpretability and post-hoc explainability results are presented along with discrimination testing results. For $g^\text{MGBM}$, intrinsic interpretability is evaluated with PD and ICE plots of monotonic prediction behavior, and post-hoc Shapley explanation analysis is used to create global and local feature importance. For $g^\text{XNN}$ intrinsic interpretability manifests as plots of sparse weights in the projection and output layers and plots of subnetwork ridge functions. Post-hoc Shapley explanation techniques are also used to generate global and local feature importance for $g^\text{XNN}$. Both $g^\text{MGBM}$ and $g^\text{XNN}$ are evaluated for discrimination using AIR, SMD, and other measures.

\subsubsection{Constrained vs. Unconstrained Model Fit Assessment}\label{ssec: c_v_uc_mort}

@@

The capability of $g^\text{XNN}$ to model nonlinear phenomenon and high-degree interactions is on display in Figure \ref{fig:mort_xnn_glob_ridge}. \ref{fig:mort_xnn_glob_ridge} \textbf{A} presents the sparse weights of the $g^\text{XNN}$ output layer, in which subnetworks 0,1,2,3,5,8, and 9 have large magnitude weights and subnetworks 4,6, and 7 have small to near-zero magnitude weights.

@@

>>>>

\subsubsection{Interpretability and Post-hoc Explanation Results}\label{ssec:int_ex_sim}

@@

Results for the mortgage data are presented in Subsections \ref{ssec:c_v_uc_mort} -- \ref{ssec:dis_mort}.  $g^\text{ANN}$ and $g^\text{XNN}$ outperform $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data, but as in Subsection \ref{ssec: c_v_uc_sim} the constrained variants of both model architectures do not show large differences in model performance with respect to unconstrained variants. Assuming that small fit differences do not outweigh the need for intrinsic model interpretability and post-hoc explainability in high-stakes, human-centered, or regulated applications, $g^\text{MGBM}$ and $g^\text{XNN}$ interpretability and post-hoc explainability results are presented along with discrimination testing results. For $g^\text{MGBM}$, intrinsic interpretability is evaluated with PD and ICE plots of monotonic prediction behavior, and post-hoc Shapley explanation analysis is used to create global and local feature importance. For $g^\text{XNN}$ intrinsic interpretability manifests as plots of sparse weights in the projection and output layers and plots of subnetwork ridge functions. Post-hoc Shapley explanation techniques are also used to generate global and local feature importance for $g^\text{XNN}$. Both $g^\text{MGBM}$ and $g^\text{XNN}$ are evaluated for discrimination using AIR, SMD, and other measures.

\subsubsection{Constrained vs. Unconstrained Model Fit Assessment}\label{ssec:c_v_uc_mort}

@@

The capability of $g^\text{XNN}$ to model nonlinear phenomenon and high-degree interactions, and to do so in an interpretable manner, is on display in Figure \ref{fig:mort_xnn_glob_ridge}. \ref{fig:mort_xnn_glob_ridge} \textbf{A} presents the sparse weights of the $g^\text{XNN}$ output layer, in which subnetworks zero, one, two, three, five, eight and nine have large magnitude weights and subnetworks four, six, and seven have small or near-zero magnitude weights. Distinctive ridge functions that feed into those weights are highlighted in \ref{fig:mort_xnn_glob_ridge} \textbf{B}. As in the subsection \ref{ssec:int_ex_sim}, ridge function plots vary with the output of the corresponding projection layer hidden unit in \ref{fig:mort_xnn_glob_ridge} \textbf{C}, and ridge functions appear to be simple functional forms that the output layer learns to combine to generate accurate predictions.

@@

\noindent \ref{fig:mort_xnn_glob_ridge} \textbf{C} displays the sparse weights of the projection layers associated with each subnetwork ridge function. For instance, subnetwork three has large weights for LTV ratio, introductory rate period flag, and introductory rate period, whereas subnetwork nine is nearly completely dominated by the weight for income. In combination, Figure \ref{fig:mort_xnn_glob_ridge} \textbf{A} presents the weighting of the ridge functions in the $g^\text{XNN}$ output, \ref{fig:mort_xnn_glob_ridge} \textbf{B} displays the ridge functions that feed directly into those output weights, and \ref{fig:mort_xnn_glob_ridge} \textbf{C} describes the weighting of multiple input features into those ridge functions.
